---
title: "Data 607 Assignment 5"
author: "Coco Donovan"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction:

I know that this assignment requires we use dplyr and tidyr, and I know that 
melt (from the reshape2 package) will be helpful in the process of combining all
of the city columns in a column for the city name and a column for the frequency.
Additionally, I will have to think of a creative solution to problem created
with not all airlines names being missing from the airline name column. The point
of this assignment is to work with the simple dataset provided, and I want to
make sure that my work would still work with a larger dataset. I essentially do
not want to be hardcoding anything, despite the fact that could be possible given
the smaller size of this data set.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading necessary packages

```{r}
library(dplyr)
library(tidyr)
library(reshape2)
```

## Name the two columns that were missing names

```{r}
messy_flights <- read.csv('flight_delays.csv')

messy_flights

messy_flights <- messy_flights %>%
  rename(Airline = X, Status = X.1, 
         'Los Angeles' = Los.Angeles, 
         'San Diego' = San.Diego,
         'San Francisco' = San.Francisco)

messy_flights
```

## Filling in missing values for airlines

```{r}
for (x in 1:nrow(messy_flights)) {
  if (messy_flights$Status[x] == 'delayed') {
    messy_flights$Airline[x] = messy_flights$Airline[x-1]
  }
}
```

## Getting rid of the blank row

```{r}
messy_flights <- messy_flights %>% drop_na()
```

## Melting the city columns together

```{r}
flights <- melt(messy_flights, flights = c(Airline, Status)) %>%
  rename(City = variable, Frequency = value)

knitr::kable(head(flights))
```

## Delays by City

```{r}
by_city <- flights %>%
  group_by(City) %>%
  summarise(percent_delayed = round(100 * sum(Frequency[Status == 'delayed'])*1.0/sum(Frequency),2), total_flights = sum(Frequency), delayed_flights = sum(Frequency[Status == 'delayed']))

knitr::kable(by_city)
```

## Delays by Airline

```{r}
by_airline <- flights %>%
  group_by(Airline) %>%
  summarise(percent_delayed = round(100 * sum(Frequency[Status == 'delayed'])*1.0/sum(Frequency),2), total_flights = sum(Frequency), delayed_flights = sum(Frequency[Status == 'delayed']))

knitr::kable(by_airline)
```

## Delays by Airline and City

```{r}
by_airline_city <- flights %>%
  group_by(Airline, City) %>%
  summarise(percent_delayed = round(100 * sum(Frequency[Status == 'delayed'])*1.0/sum(Frequency),2), total_flights = sum(Frequency), delayed_flights = sum(Frequency[Status == 'delayed'])) %>%
  arrange(City)

knitr::kable(by_airline_city)
```

## Conclusion:

From my analysis, based solely on this dataset, a few things are clear. The
first discovery I found is that San Francisco seems to experience the highest
percentages of delays on all flights. Other cities had more delayed flights, but 
in terms of percent of total San Francisco was by clearly the most delayed
airport, ALASKA experiences more delays, by percentage, than does AM WEST and 
lastly, think twice about combining AM WEST and San Francisco for your travel 
plans because there is a decent chance you will be delayed.






